---
title: "Using Fourier's transform to model forest microclimates by playing the partition of temperature harmonics"
author: "Sylvain Schmitt (F&S, CIRAD), Erwan Hingant (LAMFA, UPJV), Jonathan Lenoir (EDYSAN, CNRS)"
date: last-modified
format: 
  html:
    embed-resources: true
    toc: true
    code-fold: true
    cache: true
bibliography: references.bib
csl: mee.csl
---

```{r set}
#| message: false
# devtools::install_github("sylvainschmitt/microclimr")
library(tidyverse)
library(dataverse)
library(microclimr)
library(maps)
library(colorspace)
library(terra)
```

```{r rawdata}
#| eval: false
plots <- list(
  Aigoual = "10.48579/PRO/RSBT3C/VDNSLZ",
  Blois = "10.48579/PRO/RSBT3C/37MKWL",
  Mormal = "10.48579/PRO/RSBT3C/IJAKFT"
) |> 
  lapply(get_dataframe_by_doi, server = "data.indores.fr") %>%
  bind_rows(.id = "site")
write_tsv(plots, "data/plots.tsv")
plots %>%
  group_by(site) %>%
  summarise(lon = mean(longwgs84), lat = mean(latwgs84)) %>%
  write_tsv("data/coords.tsv")
temperatures <- list(
  Aigoual = "10.48579/PRO/ZQBQUW/MK2F35",
  Blois = "10.48579/PRO/ZQBQUW/ZL0LOC",
  Mormal = "10.48579/PRO/ZQBQUW/GBVNRR"
) %>%
  lapply(get_dataframe_by_doi, server = "data.indores.fr") %>%
  bind_rows(.id = "site")
write_tsv(temperatures, "data/hobo.tsv")
lidar <- list(
  Aigoual = "10.48579/PRO/XH62TC/P4F7FP",
  Blois = "10.48579/PRO/XH62TC/OI6WQU",
  Mormal = "10.48579/PRO/XH62TC/5RXUGU"
) %>%
  lapply(get_dataframe_by_doi, server = "data.indores.fr") %>%
  bind_rows(.id = "site")
write_tsv(lidar, "data/lidar.tsv")
lidar_sub <- lidar %>%
  filter(buffer == 5) %>%
  rename(plot = idplot) %>%
  select(site, plot, pai)
era <- read_tsv("data/era.tsv") %>%
  mutate(site = recode(dim_0, "0" = "Aigoual", "1" = "Blois", "2" = "Mormal")) %>%
  mutate(datetime = as_datetime(time)) %>%
  rename(era = tas) %>%
  select(site, datetime, era)
onf <- list.files("data/methor_data_onf/", pattern = ".txt", full.names = TRUE) %>%
  vroom::vroom(col_names = read_tsv("data/methor_data_onf/MetHor2018.org") %>%
    names()) %>%
  select(
    "Code placette",
    "Date",
    "Heure (TU)",
    "Température instantanée (°C)"
  ) %>%
  mutate(datetime = paste0(
    str_sub(Date, 7, 10), "-",
    str_sub(Date, 4, 5), "-",
    str_sub(Date, 1, 2), " ",
    str_sub(`Heure (TU)`, 1, 2), ":",
    str_sub(`Heure (TU)`, 3, 4), ":00"
  )) %>%
  mutate(datetime = as_datetime(datetime)) %>%
  rename(onf = "Température instantanée (°C)", plot = "Code placette") %>%
  filter(plot %in% c("CHP 59", "CHS 41", "HET 30")) %>%
  mutate(site = recode(plot,
    "CHP 59" = "Mormal",
    "CHS 41" = "Blois", "HET 30" = "Aigoual"
  )) %>%
  select(site, datetime, onf) %>%
  mutate(onf = gsub(",", ".", onf)) %>%
  mutate(onf = as.numeric(onf))
data <- temperatures %>%
  mutate(datetime = as_datetime(datetime)) %>%
  rename(
    hobo = t_hobo, plot = id_plot, sensor = id_sensor,
    position = position_sensor
  ) %>%
  mutate(position = recode(position, "a" = "air", "s" = "soil")) %>%
  select(site, plot, sensor, position, datetime, hobo) %>%
  left_join(era) %>%
  left_join(onf) %>%
  left_join(lidar_sub)
write_tsv(data, "data/data.tsv")
```

```{r decomposition}
#| eval: false
data <- read_tsv("data/data.tsv")
fc <- data %>%
  filter(month(datetime) %in% 5:9) %>%
  gather(
    source, temperature,
    -site, -plot, -sensor, -position, -datetime, -pai
  ) %>%
  group_by(source, site, plot, sensor, position, pai,
    datetime = floor_date(datetime, "2 hour")
  ) %>%
  summarise(temperature = mean(temperature)) %>%
  group_by(source, site, plot, sensor, position, pai) %>%
  do(fft = fft_roll(
    data = .,
    time_window = 5 * 12,
    time_col = "datetime",
    temperature_col = "temperature",
    window = "5 days",
    step = "3 days"
  )) %>%
  unnest(fft)
write_tsv(fc, "data/decomposition.tsv")
```

Aim: *Methods in Ecology and Evolution.*

## Abstract

1.  Quantifying how forest canopies buffer microclimatic temperature fluctuations is central to understanding biodiversity responses to climate change, but existing metrics (offset, slope-equilibrium) capture buffering at a single, pre-defined timescale and cannot separate daily, seasonal and interannual signals within the same temperature time series.

2.  We introduce the Discrete Fourier Transform (DFT) as a tool for microclimate ecology, decomposing paired microclimate–macroclimate temperature time series into their constituent harmonics (amplitude and phase per frequency) so that buffering can be quantified independently across timescales. We illustrate the approach with two case studies using air and soil temperature records from three French forests (Aigoual, Blois, Mormal), paired with weather-station and ERA5-Land reanalysis reference data.

3.  In the first case study, spectral decomposition isolated the dominant 24-hour cycle and showed that forest canopies buffered daily air-temperature fluctuations by 75.9–90.3% across sites, with soil buffering mean temperature more strongly than air; in the second, the same framework was used to debias ERA5-Land reanalysis data against a local weather station by comparing power spectra across frequencies, correcting both mean-temperature overestimation and underestimation of short-term variability.

4.  By resolving buffering across multiple timescales within a single, noise-robust framework and enabling principled debiasing of macroclimate reference products, the Fourier Transform offers a versatile addition to the microclimate ecologist's toolkit, complementing rather than replacing existing offset and slope-equilibrium approaches, with an accompanying open-source R package to support wider adoption.

### Keywords

microclimate buffering; Fourier transform; forest canopy; temperature time series; macroclimate debiasing; ERA5-Land; power spectrum

## Introduction

Microclimate ecology and biogeography has recently emerged as a burgeoning research field to better understand biodiversity distribution and redistribution under climate change [@kemppinen2024]. Historically, and prior to Rudolf Geiger (1894-1981) who first used the term microclimate [@geiger1965], the idea that climate can be modulated by vegetation cover was introduced by physicists, biogeographers and foresters such as Alexander von Humboldt (1769-1859), Louis Joseph Gay-Lussac (1778-1850), Antoine César Becquerel (1788-1878) and Jean-Baptiste Boussingault (1801-1887) [see @becquerel1865]. At that time, the context was very different because of the impact of massive deforestation on air quality and climate, such as in the Americas, after the arrival of European settlers.

In 1823, Boussingault travelled to Venezuela to study the drying up of the lake Valencia and confirmed von Humboldt’s intuition that forests can modify local climatic conditions across large spatial extents [@fressoz2020], by holding water locally through reduced soil evaporation and increased tree transpiration [@becquerel1865]. Boussingault concluded on a cooling effect of forests on climate, reducing mean temperature while increasing annual precipitation, suggesting that deforestation across large regions tends to warm and dry the local climate [@fuster1845; @becquerel1865; @fressoz2020]. Although neither Von Humboldt nor Boussingault mentioned the term “microclimate”, they paved the way for the early discoveries on microclimatic processes. A few decades later, in “Mémoire sur les forêts et leur influence climatérique” (1865), Becquerel reported on the calorific influence of trees on air temperatures. He recorded local air temperatures at different hours of the day, for different days of the year, between January 1863 and July 1864, and for different localities and heights near, as well as away from, isolated trees, with or without leaves. He concluded, back in 1865, that wood from trees exposed to solar and nocturnal radiations warms up and cools down the contiguous air layers surrounding trees, a feature that adds up to the non-radiative cooling effect due to tree transpiration which was already well known by the time. Almost a century later, in 1950, microclimatology became an established field, thanks to the first English edition of Rudolf Geiger’s book entitled “The Climate Near The Ground”, which percolated the field of epidemiology to better understand atmospheric air conditions in confined environments as a driver of diseases’ transmission [@sorre1961notion] \]. 

Since the mid 1980s, the advent of microclimate sensors and metrological developments have allowed meteorologists to better assess the magnitude of microclimate modulations due to vegetation cover, especially so in forest ecosystems [@defrenne2019; @defrenne2021]. The first metric used by foresters was the instantaneous difference in temperature measurements, i.e., the offset, obtained by a paired design of sensors located inside (micro) vs. outside (macro) the forest [@deFreitas1995; @carlson1997; @chen1999]. Back in 1865, Becquerel wrote “one needs to analyse temperature records made inside and outside forested areas at the same latitude, under similar geological conditions and at the same height above sea level”. In a meta-analysis compiling offset values of daily temperatures from 74 studies relying on a strictly paired design and doing exactly what Becquerel suggested in 1865, @defrenne2019 have found that mean and maximum sub-canopy temperatures were, on average, cooler than temperatures outside forests by 1.7 ± 0.3°C and 4.1 ± 0.5 °C, respectively, a pattern which was even more pronounced in tropical forests compared to temperate or boreal forests. Another metric often used to assess microclimate modulations is the slope parameter of a linear model linking paired time series of hourly temperature records from a microclimate sensor located beneath forest canopies (micro) versus a neighboring weather station located outside the forest (macro), such that slope values lower and greater than 1 suggest buffering and amplifying effects of temperature fluctuations over time, respectively [@gril2023]. However, none of these two metrics allow for a direct decomposition of the periodic signal of temperature fluctuations over a long time series, in order to isolate daily fluctuations from seasonal variations and interannual trends. Here, we propose to fill this gap by using the Fourier Transform (FT) on paired temperature time series, micro vs. macro, to decompose each temperature time series in a sum of its harmonics.

Each harmonic corresponds to a wave (cyclic) signal with its own range of variation (amplitude) and its own periodicity (frequency) to fulfill a complete cycle. Thanks to FT, any cyclic signal, such as temperature, can be written as a periodic function summing an infinity of harmonics and thus can be fully reconstructed by virtue of the knowledge of the ordered collection (vector) of amplitude values, called power spectrum, and its corresponding vector of phase values, i.e., time shifts relative to the reference, called Fourier decomposition of the signal. By extracting the power spectrum of both the microclimatic and macroclimatic temperature time series, it is then possible to analyse the vector of ratios of amplitude values (micro/macro), which corresponds to a full decomposition of microclimate modulations for increasing frequencies, such that ratio values lower and greater than 1 mean buffering and amplifying effects, respectively. It is then possible to isolate microclimate modulations for the frequency of interest (e.g., daily) to detrend the signal, i.e., ignoring lower-frequency signals (e.g., seasonal and interannual), and to remove potential noise due to the harmonics with very low amplitudes in the harmonic series. 

To illustrate the relevance of FT for the field of microclimate ecology and biogeography, we present two case studies. First, we use FT to assess the buffering capacity of temperate deciduous forest ecosystems during the vegetation season. We rely on an existing dataset of air and soil temperature time series from a network of microclimate loggers installed at 1-m height aboveground and 8-cm depth belowground, respectively, and beneath the canopy of three state forests in France [@gril2023; @gril2024; @gril2025]. This dataset is paired with a set of time series representing the reference temperature (macroclimate) extracted from either the ERA5-Land global reanalysis [@muñoz-sabater2021] or three weather stations located outside but nearby the three studied state forests. Second, because the choice of the macroclimatic reference matters to assess microclimatic modulations, we apply FT to debias a temperature time series from the ERA5-Land reanalysis at a local site where a paired time series from a ground-truth weather station is available so that this debiasing procedure can be implemented on another location where weather station data are unavailable.

## Method: Spectral Decomposition of Temperature Time Series via the Fourier Transform

Temperature records measured continuously over time — whether by a microclimate sensor inside a forest or by a macroclimate reanalysis product or weather station in open field — are inherently periodic signals. Daily heating and cooling, seasonal rhythms, and longer climatic cycles all superimpose upon one another to produce the complex thermal fluctuations that ecologists observe. To disentangle these overlapping rhythms and quantify how the forest modifies each of them independently, we decomposed temperature time series using the Discrete Fourier Transform (DFT).

**Mathematical framework.** Let $T(t_j)$, $j = 0, 1, \ldots, N-1$, denote a temperature time series sampled at regular intervals $\Delta t$ (here, hourly) over a total duration $N \Delta t$. The DFT decomposes this signal into a sum of complex sinusoids:

$$\hat{T}(f_k) = \sum_{j=0}^{N-1} T(t_j)\, e^{-2\pi i\, j\, k / N}, \quad k = 0, 1, \ldots, N-1,$$

where $f_k = k / (N \Delta t)$ is the $k$-th frequency (in cycles per unit time) and $i$ is the imaginary unit. The amplitude spectrum $|\hat{T}(f_k)|$ quantifies the contribution of each frequency to the total signal, while the phase $\arg(\hat{T}(f_k))$ captures the timing of each periodic component. In practice, computations are performed via the Fast Fourier Transform (FFT) algorithm, which reduces the computational complexity from $\mathcal{O}(N^2)$ to $\mathcal{O}(N \log N)$.

The inverse DFT reconstructs the original signal exactly:

$$T(t_j) = \frac{1}{N} \sum_{k=0}^{N-1} \hat{T}(f_k)\, e^{2\pi i\, j\, k / N}.$$

Because temperature is a real-valued signal, the spectrum is Hermitian symmetric ($\hat{T}(f_{N-k}) = \overline{\hat{T}(f_k)}$), so only the first $N/2$ frequencies carry independent information. We therefore report the one-sided amplitude spectrum scaled by $2/N$ (except at the zero-frequency and Nyquist components).

**Ecological interpretation.** Key frequencies in a temperature record correspond to ecologically meaningful cycles, for instance the diurnal cycle with a frequency of 1/24 hr (@fig-proj). Expressing temperature variability in the frequency domain thus allows us to ask not simply "how much cooler is the forest interior?" but rather "at which timescales does the forest attenuate or amplify climatic variability?" Forest canopies are expected to buffer high-frequency (diurnal) fluctuations more strongly than low-frequency (seasonal) ones, because the physical mechanisms responsible — thermal inertia of soil and litter, latent heat exchanges, and radiative shading — operate preferentially at short timescales. As illustrated in @fig-proj for a representative 5-day temperature time series, the spectral decomposition successfully resolves the signal into its constituent frequencies: a mean microclimate temperature of 16.24°C, a diurnal amplitude of 3°C at the 24-hour period, and higher harmonics — together revealing that the forest microclimate substantially buffers the diurnal temperature extremes relative to the macroclimate.

```{r proj_freq}
#| label: fig-proj
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 8
#| fig-cap: "Temperature time series decomposition, using Fourier’s transform, and reprojection for selected frequencies. We used an example of two paired (macroclimate vs. microclimate) hourly temperature time series from August 7th to August 12th 2023 (upper-left panel). The microclimate (Micro) temperature time series was extracted from one microclimate logger (HOBO, dark lines) installed in plot #59_58 in the Mormal forest and recording sub-canopy air temperature at 1-m height above ground. The macroclimate (Macro) temperature time series was extracted from one neighboring weather station located in an open landscape near the forest and managed by the “Office National des Forêts” (ONF, light lines): it follows the standards of the World Meteorological Organisation (WMO). The lower left panel represents the resulting power spectrum after Fourrier’s decomposition applied on the 6-days long microclimate (HOBO, dark bars) and macroclimate (ONF, light bars) time series. Frequencies above 1/24h are highlighted in blue (“trend”) while the frequency at 1/24h and its corresponding harmonics are highlighted in red (“day”). The resulting reconstruction of temperature time series for the highlighted frequencies are given in the upper right panel (overall trend in blue) and lower right panel (daily trend in red)."
Sys.setlocale("LC_TIME", "en_US.UTF-8") %>%
  invisible()
cols <- c("darkblue", "firebrick", "darkgrey")
data <- "data/data.tsv" %>%
  read_tsv() %>%
  filter(
    year(datetime) == 2023,
    month(datetime) == 8
  ) %>%
  filter(
    position == "air",
    plot == "59_58"
  ) %>%
  arrange(datetime) %>%
  filter(day(datetime) %in% 7:12)
fc_h <- fft_tab(data$hobo, 24 * 6) %>%
  mutate(coeff_sup = ifelse(frequency > 1 / (24 + 1), 0,
    fft_coefficient(amplitude, phase)
  )) %>%
  mutate(type = ifelse(frequency > 1 / (24 + 1), "other", "Overall trend")) %>%
  mutate(coeff_day = ifelse(frequency %in% c((1:12) / 24),
    fft_coefficient(amplitude, phase), 0
  )) %>%
  mutate(type = ifelse(frequency %in% c((1:12) / 24), "Daily fluctuations", type))
data$t_h_sup <- fft_reconstruct(fc_h$coeff_sup, fc_h$frequency[-1], 1:144)
data$t_h_day <- fft_reconstruct(fc_h$coeff_day, fc_h$frequency[-1], 1:144) + fc_h$power[1]
fc_o <- fft_tab(data$onf, 24 * 6) %>%
  mutate(coeff_sup = ifelse(frequency > 1 / (24 + 1), 0,
    fft_coefficient(amplitude, phase)
  )) %>%
  mutate(type = ifelse(frequency > 1 / (24 + 1), "other", "Overall trend")) %>%
  mutate(coeff_day = ifelse(frequency %in% c((1:12) / 24),
    fft_coefficient(amplitude, phase), 0
  )) %>%
  mutate(type = ifelse(frequency %in% c((1:12) / 24), "Daily fluctuations", type))
data$t_o_sup <- fft_reconstruct(fc_o$coeff_sup, fc_o$frequency[-1], 1:144)
data$t_o_day <- fft_reconstruct(fc_o$coeff_day, fc_o$frequency[-1], 1:144) + fc_o$power[1]
g_data <- ggplot(data, aes(x = datetime)) +
  geom_line(aes(y = era, col = "Macro")) +
  geom_line(aes(y = hobo, col = "Micro")) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  scale_color_manual("", values = c(lighten("black", 0.5), "black")) +
  xlab("Time (day)") +
  ylab("Temperature [°C]") +
  ylim(8, 28) +
  theme(
    legend.position = c(0.3, 0.8),
    legend.background = element_blank(),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 12)
  )
g_fc <- fc_h %>%
  left_join(fc_o,
    by = join_by(frequency, period, type),
    suffix = c("", "_onf")
  ) %>%
  filter(period != 0) %>%
  ggplot(aes(frequency, power, fill = type)) +
  geom_col(aes(y = power_onf), alpha = .5) +
  geom_col() +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  scale_x_continuous(
    breaks = c(1 / (24 * 5), 1 / 24, 1 / 12, 1 / 8, 1 / 6),
    labels = c("5d", "24", "12", "8", "6")
  ) +
  scale_fill_manual("", values = cols, breaks = c(
    "Overall trend",
    "Daily fluctuations"
  )) +
  theme(legend.position = c(0.8, 0.6)) +
  xlab("Period [hr]") +
  ylab("Power [°C]") +
  theme(
    axis.text.x = element_text(angle = 0, size = 12),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12)
  ) +
  annotate("text",
    x = 3 / 24, y = 4.5,
    label = paste0("Mean:", round(fc_h$power[1], 2)), size = 5
  ) +
  annotate("text",
    x = 3 / 24, y = 5, alpha = .5,
    label = paste0("Mean:", round(fc_o$power[1], 2)), size = 5
  )
g_sup <- ggplot(data, aes(datetime)) +
  geom_line(aes(y = t_o_sup), col = lighten(cols[1], 0.5)) +
  geom_line(aes(y = t_h_sup), col = cols[1]) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  xlab("Time (day)") +
  ylab("") +
  ylim(8, 28) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold")
  ) +
  annotate("text",
    x = data$datetime[10], y = 25, hjust = 0,
    label = "Overall trend", col = cols[1], size = 5.5, fontface = "bold"
  )
g_day <- ggplot(data, aes(datetime)) +
  geom_line(aes(y = t_o_day), col = lighten(cols[2], 0.5)) +
  geom_line(aes(y = t_h_day), col = cols[2]) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  xlab("Time (day)") +
  ylab("") +
  ylim(8, 28) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold")
  ) +
  annotate("text",
    x = data$datetime[10], y = 25, hjust = 0,
    label = "Daily fluctuations", col = cols[2], size = 5.5, fontface = "bold"
  )
g <- cowplot::plot_grid(g_data, g_sup, g_fc, g_day,
  rel_widths = c(1.5, 1), align = "v"
)
cowplot::save_plot("figs/fourier.png",
  plot = g,
  base_width = 10,
  base_height = 8,
  dpi = 300,
  bg = "white"
)
g
```

```{r ft_gif}
#| eval: false
library(gganimate)
data <- "data/data.tsv" %>%
  read_tsv() %>%
  filter(
    year(datetime) == 2023,
    month(datetime) == 8
  ) %>%
  filter(
    position == "air",
    plot == "59_58"
  ) %>%
  arrange(datetime) %>%
  filter(day(datetime) %in% 7:12)
fc_h <- fft_tab(data$hobo, 24 * 6) %>%
  mutate(n = 0:(n() - 1))
fc_o <- fft_tab(data$onf, 24 * 6) %>%
  mutate(n = 0:(n() - 1))
proj <- function(i, fc_h, fc_o, data) {
  if (i != 0) {
    fc_h <- fc_h %>%
      mutate(coefficient = ifelse(n > i, 0, fft_coefficient(amplitude, phase)))
    fc_o <- fc_o %>%
      mutate(coefficient = ifelse(n > i, 0, fft_coefficient(amplitude, phase)))
    data$t_h <- fft_reconstruct(fc_h$coefficient, fc_h$frequency[-1], 1:144)
    data$t_o <- fft_reconstruct(fc_o$coefficient, fc_o$frequency[-1], 1:144)
    return(data %>% mutate(n = i))
  } else {
    data$t_h <- fc_h$power[1]
    data$t_o <- fc_o$power[1]
    return(data %>% mutate(n = i))
  }
}
|> 
g <- lapply(0:max(fc_h$n), proj, fc_h, fc_o, data) %>%
  bind_rows() %>%
  ggplot(aes(x = datetime)) +
  geom_line(aes(y = t_o, col = "Macro")) +
  geom_line(aes(y = t_h, col = "Micro")) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  scale_color_manual("", values = c(lighten("black", 0.5), "black")) +
  xlab("Time (day)") +
  ylab("Temperature [°C]") +
  ylim(8, 28) +
  theme(
    legend.position = "none",
    legend.background = element_blank(),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 12)
  ) +
  transition_states(n, state_length = 0, transition_length = c(72:0 + 1) * 2) +
  labs(title = "N≤{as.integer(closest_state)}")
animate(g, renderer = gifski_renderer(width = 3000, height = 1800))
anim_save("fourier.gif")
```

## Study sites and data

*Imprint project description* [@gril2024] (@fig-sites)*.* Additionally macroclimatic data were retrieved either from ERA5-Land global reanalysis [@muñoz-sabater2021] or from ONF ... *to be further described.*

```{r sites}
#| label: fig-sites
#| message: false
#| warning: false
#| fig-cap: "Location, across the three study sites in France, of the macroclimatic and microclimatic temperature time series used in the two study cases. The three study sites are the Mormal, Blois and Aigoual forests located in the northern, central and southern part of France (upper-right panel). The Aigoual forest (upper-right panel) is a mountain forest under Mediterranean climate influences. The Blois forest (lower-left panel) is a lowland and periurban forest under oceanic climate influences. The Mormal forest (lower-right panel) is a lowland and rural forest under continental climate influences. The raster grid depicted in the upper-left panel shows macroclimatic temperature conditions from the ERA5-Land reanalysis at 9-km resolution for the August 11th 2023 at 12pm. Points coloured according to ERA5-Land temperature conditions show the location of microclimate loggers (HOBO) in the field while white stars show the location of the ONF weather stations."
#| fig-width: 10
#| fig-height: 8
sites <- read_tsv("data/plots.tsv") %>%
  group_by(site) %>%
  summarise_all(mean) %>%
  rename(long = longwgs84, lat = latwgs84)
stations <- readxl::read_xlsx("data/methor_data_onf/coord_stations_meteo_renecofor.xlsx") %>%
  mutate(site = c("Mormal", "Blois", "Aigoual")) %>%
  mutate(lon = X, lat = Y)
france <- map_data("france")
fr <- ggplot(france, aes(x = long, y = lat)) +
  tidyterra::geom_spatraster(data = rast("data/erafrance.nc"), alpha = .9) +
  geom_point(data = sites, size = 4, col = "black") +
  ggrepel::geom_text_repel(
    data = sites, aes(label = site), size = 6,
    color = "white", fontface = "bold",
    bg.color = "black", bg.r = 0.15, force = 3,
    max.overlaps = Inf
  ) +
  coord_sf(expand = FALSE) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  xlab("Longitude") +
  ylab("Latitude") +
  theme(
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 12, face = "bold"),
    plot.background = element_rect(fill = "white", color = NA),
    aspect.ratio = 1
  ) +
  scico::scale_fill_scico(
    name = "°C",
    palette = "lajolla",
    na.value = "white", direction = -1
  ) +
  theme(
    legend.text = element_text(size = 11),
    legend.title = element_text(size = 12, face = "bold")
  )
si <- read_tsv("data/plots.tsv") %>%
  rename(plot = idplot) %>%
  left_join(read_tsv("data/data.tsv") %>%
    filter(
      position == "air",
      datetime == as_datetime("2023-08-11 12:00:00")
    ) %>%
    select(plot, hobo)) %>%
  rename(lon = longwgs84, lat = latwgs84) %>%
  group_by(site) %>%
  mutate(hobo = ifelse(is.na(hobo), mean(hobo, na.rm = TRUE), hobo)) %>%
  split(.$site) %>%
  lapply(function(x) {
    points <- sf::st_as_sf(x, coords = c("lon", "lat"), crs = 4326)
    lat_scale <- cos(mean(x$lat) * pi / 180)
    bbox <- sf::st_bbox(points)
    cx <- (bbox["xmin"] + bbox["xmax"]) / 2
    cy <- (bbox["ymin"] + bbox["ymax"]) / 2
    half_lat <- max(
      (bbox["xmax"] - bbox["xmin"]) * lat_scale,
      bbox["ymax"] - bbox["ymin"]
    ) / 2 + 0.02
    sq_bbox <- c(
      xmin = cx - half_lat / lat_scale, xmax = cx + half_lat / lat_scale,
      ymin = cy - half_lat, ymax = cy + half_lat
    )
    names(sq_bbox) <- c("xmin", "xmax", "ymin", "ymax")
    sq_sfc <- sf::st_bbox(sq_bbox, crs = 4326) %>% sf::st_as_sfc()
    bg <- maptiles::get_tiles(sq_sfc, provider = "Esri.WorldImagery", zoom = 12)
    ggplot(points) +
      tidyterra::geom_spatraster_rgb(data = bg, alpha = .9) +
      geom_sf(aes(fill = hobo), shape = 21, col = "black", size = 4) +
      geom_text(
        data = filter(stations, site == unique(x$site)),
        aes(x = lon, y = lat),
        label = "★", color = "white", size = 6, fontface = "bold"
      ) +
      annotate("text",
        x = Inf, y = Inf, label = unique(x$site),
        hjust = 1.1, vjust = 1.8,
        size = 6, color = "white", fontface = "bold",
        bg.color = "black", bg.r = 0.15, force = 3
      ) +
      scale_x_continuous(breaks = scales::extended_breaks(n = 5), labels = function(x) paste0(format(x, nsmall = 1), "°E")) +
      scale_y_continuous(breaks = scales::extended_breaks(n = 5), labels = function(x) paste0(format(x, nsmall = 1), "°N")) +
      coord_sf(
        xlim   = c(cx - half_lat / lat_scale, cx + half_lat / lat_scale),
        ylim   = c(cy - half_lat, cy + half_lat),
        expand = FALSE
      ) +
      theme_bw() +
      theme(
        axis.text = element_text(size = 11),
        axis.title = element_text(size = 12, face = "bold"),
        aspect.ratio = 1,
        panel.grid = element_blank()
      ) +
      scico::scale_fill_scico(
        palette = "lajolla",
        name = "°C",
        na.value = "darkgrey", direction = -1
      ) +
      xlab("Longitude") +
      ylab("Latitude") +
      theme(
        legend.text = element_text(size = 11),
        legend.title = element_text(size = 12, face = "bold")
      )
  })
g <- cowplot::plot_grid(plotlist = c(fr, si), align = "hv")
cowplot::save_plot("figs/sites.png",
  plot = g,
  base_width = 10,
  base_height = 8,
  dpi = 300,
  bg = "white"
)
g
```

## Study case 1 - **Forest microclimate assessment**

We decomposed the temperature series for the Aigoual, Blois and Mormal locations from the macroclimate reference products. These products were either from local ONF weather stations on open land or from the corresponding pixel in the ERA5-Land reanalysis. We also decomposed the temperature series from HOBO microclimate sensors, which recorded either air or soil temperatures. Our analysis focused on the leaf-on season, when a visible forest microclimate effect was present, from May to September. We used Fourier transform decomposition with a sliding window of five days and a shift of three days, analysing only complete data series. Based on this decomposition, we created power spectra for air temperatures (see @fig-ps), which highlight the power of temperature variation (°C) per period or frequency (here, period in hours; see @fig-proj for an explanation).

We found that the mean air temperature was similar for the ONF macroclimates and the air microclimates (15.4 °C, 19.3 °C and 16.3 °C for Aigoual, Blois and Mormal, respectively). However, the ERA5-Land dataset overestimated the mean temperatures (17.4 °C, 20.3 °C and 17.3 °C), and the mean temperature of the soil microclimate was lower than that of the macroclimate (13.3 °C, 17.2 °C and 15.2 °C). The power spectrum revealed most variations to occur at a 24-hour daily variation and its harmonics. 

We then computed the microclimate buffers as the ratio of mean temperature and power for the 24-hour period (@fig-ratios) for the HOBO microclimate sensors, compared to local ONF weather stations on open land. This was because the power spectrum revealed discrepancies in the ERA5-Land data, which would bias the buffer estimates. Our analysis focused on sensors with at least two-thirds of the time series with full data. We found that, in the three sites, soil buffers had a stronger effect on mean temperatures than air buffers. Air buffers did not change significantly the mean temperature, but they reduced the daily 24-hour temperature fluctuations, decreasing night-time temperatures and increasing daytime temperatures by 90.3%, 86.1% and 75.9% for Aigoual, Blois and Mormal, respectively.

We also obtained the results of @gril2024, by exploring all the buffers per plot (see @fig-buffer) and linking them to the plot vegetation. The plant area index explained up to 67% of the variation in microclimate air temperature across the buffers (see @fig-pai). However, our methodology enabled us to explore the multi-temporal variation of the buffers in more detail with several periods and frequencies, and to integrate potential dephasing (see discussion).

```{r ps}
#| label: fig-ps
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 8
#| fig-cap: Power spectrum for air and soil temperatures, using HOBO microclimate temperatures and ERA5-Land and ONF macroclimate temperatures, for the locations of Aigoual, Blois and Mormal
t <- "data/decomposition.tsv" %>%
  read_tsv() %>%
  group_by(source, site, position, frequency, period) %>%
  summarise(
    sd = sd(power, na.rm = TRUE),
    power = mean(power, na.rm = TRUE)
  ) %>%
  mutate(type = ifelse(source == "hobo", "Micro", "Macro")) %>%
  mutate(source = recode(source,
    "era" = "ERA5-Land",
    "onf" = "ONF weather station",
    "hobo" = paste("HOBO", position)
  )) %>%
  filter(!(type == "Macro" && position == "soil"))
g <- t %>%
  filter(period != 0) %>%
  ggplot(aes(frequency, power)) +
  geom_col(fill = "darkgrey") +
  geom_linerange(aes(ymin = power - sd, ymax = power + sd)) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  xlab("Period [hr]") +
  ylab("Power [°C]") +
  scale_x_continuous(
    breaks = c(1 / 120 * 2, 1 / 24 * 2, 1 / 12 * 2, 1 / 8 * 2, 1 / 6 * 2, 1 / 3 * 2),
    labels = c("5d", "24", "12", "8", "6", "3")
  ) +
  theme(
    legend.position.inside = c(0.8, 0.8),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold"),
    strip.text = element_text(size = 12, face = "bold")
  ) +
  facet_grid(site ~ paste0(type, "\n", source)) +
  geom_text(
    x = 1 / 3, y = 6,
    aes(label = paste0("Mean: ", round(power), "±", round(sd))),
    data = filter(t, period == 0), size = 5
  )
cowplot::save_plot("figs/buffer.png",
  plot = g,
  base_width = 10,
  base_height = 8,
  dpi = 300,
  bg = "white"
)
g
```

We then computed the microclimate buffers as ratios of mean temperature and power for the 24-hr period (@fig-ratios). We look only at sensors with at least 2/3 of expected windows (160). *Interperation.*

```{r ratios}
#| label: fig-ratios
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 8
#| fig-cap: Mean temperature vs. 24-h amplitude ratios of HOBO microclimate sensors against ONF weather station for air and soil sensors at Aigoual, Blois and Mormal forests. The point represents the median, the thick line the 50 percent distribution and the thine line the 90 percent distribution.
g <- read_tsv("data/decomposition.tsv") %>%
  filter(source != "era") %>%
  filter(period %in% c(0, 12)) %>%
  select(-amplitude, -phase, -frequency) %>%
  pivot_wider(names_from = source, values_from = power) %>%
  na.omit() %>%
  mutate(ratio = hobo / onf) %>%
  select(-hobo, -onf) %>%
  mutate(period = paste0("period_", period)) %>%
  group_by(position, site, period) %>%
  summarise(
    q05 = quantile(ratio, .05),
    q25 = quantile(ratio, .25),
    q5 = quantile(ratio, .5),
    q75 = quantile(ratio, .75),
    q95 = quantile(ratio, .95)
  ) %>%
  gather(metric, value, -position, -site, -period) %>%
  mutate(period_metric = paste0(period, "_", metric)) %>%
  select(-period, -metric) %>%
  pivot_wider(names_from = period_metric, values_from = value) %>%
  ggplot(aes(period_0_q5, period_12_q5, col = position)) +
  geom_vline(xintercept = 1, col = "darkgrey") +
  geom_hline(yintercept = 1, col = "darkgrey") +
  geom_segment(aes(x = period_0_q25, xend = period_0_q75)) +
  geom_segment(aes(x = period_0_q05, xend = period_0_q95), linewidth = .2) +
  geom_segment(aes(y = period_12_q25, yend = period_12_q75)) +
  geom_segment(aes(y = period_12_q05, yend = period_12_q95), linewidth = .2) +
  geom_point() +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  ggrepel::geom_text_repel(aes(label = site), col = "black", size = 4.5) +
  scale_color_manual("", values = c("#a9d9e5", "#c89867")) +
  xlab("Mean temperature ratio [°C]") +
  ylab("24-h amplitude ratio [°C]") +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12)
  )
cowplot::save_plot("figs/buffer2.png",
  plot = g,
  base_width = 10,
  base_height = 8,
  dpi = 300,
  bg = "white"
)
g
```

## Study case 2 - Macroclimate debiasing

Previous reanalyses of microclimate buffers in the Aigoual, Blois and Mormal forests clearly demonstrated the importance of considering the macroclimate when conducting microclimate studies [@hes2025]. Field data from open-air standard stations, such as the ONF station used here, are not always available for our study site of interest, and may be incomplete or only available for a distant site. Conversely, global climate data reanalyses such as ERA5-Land [@muñoz-sabater2021] are available worldwide and cover a long time span, but they include biases, such as overestimation of mean temperature and underestimation of fine temporal variations due to the reanalysis smoothing effect (see @fig-ps).  In this case, Fourier transform decomposition could help us to bridge the gap between local field data from open-air standard stations at a nearby site and ERA5-Land reanalysis data, enabling us to obtain a qualitative, debiased reference macroclimate product for use in microclimate studies (@fig-debias).

Typically, studies may lack local field reference data for a given site, but have similar data from a nearby site. In each case, the ERA5-Land reanalysis will have data for the corresponding pixels, and the bias is likely to be consistent across the nearby sites. The debiasing scheme in this case uses the local station time series and ERA5-Land data from the nearby site to create a debiasing model based on the Fourier transform. This model is then applied to the ERA5-Land data for the site of interest to produce a debiased temperature series for the study site. This series can be used for microclimate studies, modelling, and other ecological applications. For example, we applied this to the temperature time series of the Mormal forest for the five days around 7 August 2023. Using power spectrum decomposition (see @fig-debias), we found a higher mean temperature in the ERA5-Land data (17.51°C compared to 17.01°C), as well as lower 24-hour period variation and associated harmonics (3.96°C compared to 5.38°C). We then used the decomposition for each period and frequency to calculate the ratio of the powers of ERA5-Land and the local ONF station, which was used to debias the time series at the site of interest. Globally, we found lower power values for the local ONF station compared to ERA5-Land (positive ratios), and higher discrepancies in power for smaller frequencies and longer periods compared to higher frequencies and shorter periods (regression line in blue in the bottom right panel of @fig-debias).

The advantage of such debiasing is that, by its very nature, it provides a better account of the multitemporal variation in temperature series than methodologies that work on a specific timescale, such as yearly, monthly or daily averages (see discussion).

```{r debias}
#| label: fig-debias
#| message: false
#| warning: false
#| fig-cap: "Debiasing of the ERA5-Land reanalysis for a site without a local station. This is based on data from a nearby site and Fourier's transform. Top left: the debiasing scheme uses a nearby site with a local station and ERA5-Land data to create a debiasing model based on the Fourier transform. This model is then applied to the ERA5-Land data for the site of interest, producing a debiased temperature series for the study site. This series can be used for microclimate studies, modelling and other ecological applications. Top right: the ERA5-Land and ONF temperature series for a nearby site over a five-day period. Bottom left: the resulting power spectrum, which highlights, among other things, the higher mean temperature of ERA5-Land and the lower 24-hour period variation and associated harmonics. Bottom right: the pairwise ratios between the ONF weather station and the ERA5-Land reanalysis for all periods and frequencies, which are used as a debias function. Note that we are using a log scale. The decrease with frequency indicates that the differences are smaller for higher frequencies (i.e. shorter time periods)."
#| fig-width: 10
#| fig-height: 8
cols <- c("#0F6E56", "#BA7517")
data <- "data/data.tsv" %>%
  read_tsv() %>%
  filter(
    year(datetime) == 2023,
    month(datetime) == 8
  ) %>%
  filter(
    position == "air",
    plot == "59_58"
  ) %>%
  arrange(datetime) %>%
  filter(day(datetime) %in% 7:12)
g_data <- ggplot(data, aes(x = datetime)) +
  geom_line(aes(y = era, col = "ERA5-Land")) +
  geom_line(aes(y = onf, col = "ONF")) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  scale_color_manual("", values = cols) +
  xlab("") +
  ylab("Temperature [°C]") +
  ylim(8, 28) +
  theme(
    legend.position = c(0.3, 0.8),
    legend.background = element_blank(),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 12)
  )
fc_e <- fft_tab(data$era, 24 * 6)
fc_o <- fft_tab(data$onf, 24 * 6)
g_fc <- list(era = fc_e, onf = fc_o) %>%
  bind_rows(.id = "origin") %>%
  filter(period != 0) %>%
  ggplot(aes(frequency, power, fill = origin)) +
  geom_col(position = "dodge") +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  scale_x_continuous(
    breaks = c(1 / (24 * 5), 1 / 24, 1 / 12, 1 / 8, 1 / 6),
    labels = c("5d", "24", "12", "8", "6")
  ) +
  theme(
    legend.position = c(0.8, 0.6),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12)
  ) +
  xlab("Period [hr]") +
  ylab("Power [°C]") +
  scale_fill_manual(guide = "none", values = cols) +
  annotate("text",
    x = 5 / 24, y = 3,
    label = paste0(
      "Mean:", round(fc_e$power[1], 2), " vs. ",
      round(fc_o$power[1], 2)
    ), size = 5
  )
g_cor <- list(era = fc_e, onf = fc_o) %>%
  bind_rows(.id = "origin") %>%
  select(origin, frequency, period, amplitude) %>%
  pivot_wider(names_from = origin, values_from = amplitude) %>%
  ggplot(aes(frequency, onf / era)) +
  geom_hline(yintercept = 1, col = "darkgrey") +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point() +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  scale_x_continuous(
    breaks = c(1 / (24 * 5), 1 / 24, 1 / 12, 1 / 8, 1 / 6),
    labels = c("5d", "24", "12", "8", "6")
  ) +
  scale_y_log10() +
  xlab("Period [hr]") +
  ylab("Ratio ONF vs. ERA5-Land") +
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold")
  )
svg_raster <- rsvg::rsvg_nativeraster("data/debiasing_scheme_v2.svg")
g_scheme <- grid::rasterGrob(svg_raster, interpolate = TRUE)
g <- cowplot::plot_grid(g_scheme, g_data, g_fc, g_cor)
cowplot::save_plot("figs/bias.png",
  plot = g,
  base_width = 10,
  base_height = 8,
  dpi = 300,
  bg = "white"
)
g
```

## Discussion

Here, we show that the Discrete Fourier Transform offers a powerful and largely untapped tool for microclimate ecology and biogeography, extending beyond the offset and slope-equilibrium approaches traditionally used to quantify canopy buffering. By decomposing temperature time series into their constituent frequencies, the Fourier Transform captures buffering dynamics across multiple timescales simultaneously, is inherently robust to sensor noise, and yields a novel, physically grounded metric of buffering intensity based on power spectrum variance decomposition. Beyond microclimate buffering itself, we further demonstrate that the same framework can be used to debias macroclimate reference products against local observations. Together, these results position the Fourier Transform as a versatile and complementary, rather than competing, addition to the microclimate ecologist's toolkit.

### **Microclimate buffer amplitude analyses**

Using spectral decomposition of temperature time series via the Discrete Fourier Transform, we were able to extend previous analyses of microclimate buffering beyond the offset approach [@deFreitas1995] or the slope-equilibrium approach [@gril2023]. Specifically, the Fourier Transform allowed: (i) a finer analysis of multi-temporal variation in buffering, capturing several periods and frequencies simultaneously; (ii) a more robust methodology for dealing with measurement and sensor noise; and (iii) a new metric, based on power spectrum variance decomposition, to quantify the proportion of climatic variability buffered by forest canopies.

Indeed, the offset [@deFreitas1995] and slope-equilibrium [@gril2023] approaches both require defining the timescale(s) to work with beforehand. Previous studies have worked with hourly [@gril2024], daily [@defrenne2019], monthly [@hes2025], seasonal or yearly [@gril2025] statistics, and have rarely combined more than two timescales [@hes2025] or operated at sub-daily resolution [but see @pinheiro2011]. We believe that the Fourier Transform allows a finer analysis integrating all these timescales and others within a single approach, depending on how the sliding time window of the analysis is constructed (discussed below). Moreover, the Fourier Transform is not in opposition to the offset and slope-equilibrium approaches: once the power spectrum is built for each frequency/period, the microclimate buffer still needs to be computed, for instance using an offset as we did in study case 1 (@fig-ratios), though other approaches could be considered.

In addition, the offset and slope-equilibrium approaches are more sensitive to outliers arising from sensor malfunction, often requiring a preliminary data-cleaning stage to remove such artefacts. For instance, [@hes2025] discarded points falling more than five standard deviations from the mean over a monthly rolling window. In the time series shown in Figure 1, such noise is clearly visible at the start of August 9th, where a malfunctioning microclimate sensor produces a temperature drop relative to the macroclimate signal. Interestingly, this artefact is not captured in either the overall trend or the daily fluctuations, since such small-scale variations are only reflected in the high-frequency (low-period) component of the signal making the Fourier Transform robust to this type of noise not visible in the lower-frequency (higher-period) components used in subsequent analyses.

Last but not least, based on Parseval's Theorem, the Fourier Transform provides a further metric: by decomposing the variance of the power spectrum, one can compute the total "energy" of the microclimate signal relative to the macroclimate signal. In our case, an energy ratio of 86% indicates that the forest microclimate dissipates part of the variation present in the macroclimate.

### **The advantages of Fourier for dephasing and debiasing**

Beyond the extraction of multi-frequency amplitudes from temperature time series, the Fourier Transform also allows the extraction of signal phases, providing a better account of dephasing between signals across timescales. For instance, depending on the temporal resolution considered, soil temperature changes lag behind air temperature at short timescales owing to differences in thermal properties between media [@yin2024]. In such cases, the offset and slope-equilibrium approaches introduce statistical noise, since the temperature measurements are no longer correctly time-matched: part of the variation around the offset, and around the slope-equilibrium regression, then reflects this time lag rather than microclimate buffering itself. By explicitly accounting for dephasing, the Fourier Transform resolves this issue (results not shown). We did not focus on the phase component of the decomposition here, but future studies could exploit it for phase-shift analysis, interpreting temporal offsets between macroclimate and microclimate signals as indicators of thermal inertia and canopy lag [e.g. @bush2016].

We first applied the Fourier Transform to microclimate assessment, as shown in study case 1. A further, less commonly discussed problem in microclimatology, and in applied climatology more broadly, is how to obtain a robust reference macroclimate product against which to estimate microclimate buffering. For instance, [@hes2025] corrected ERA5-Land reanalysis products in French Guiana against local MétéoFrance weather stations using a mean-and-variance correction, requiring separate corrections for each temporal frequency considered (hourly and monthly). We believe the Fourier Transform offers a clear advantage here for debiasing global macroclimate products against local data, as demonstrated in study case 2: it provides a fast, simple correction that remains robust across multiple timescales simultaneously. This approach is already used in several other domains: for example to correct CMIP6 simulations in climatology [@kusumastuti2022], to improve downscaling [@utami2024], or in hydrology [@kendall2007], and we encourage its wider adoption in ecology and biogeography.

### **Limitations**

Like any approach, the Fourier Transform has limitations. Compared to the offset [@deFreitas1995] or slope-equilibrium [@gril2023] approaches, which are increasingly adopted across a wide range of ecosystems and contexts, our approach still requires testing for scalability to other forest types and climates, in order to evaluate whether the Fourier Transform approach generalises beyond our three French study sites (Aigoual, Blois, Mormal). Methodologically, there is no specific reason for the Fourier Transform to be restricted to these cases, though this remains to be demonstrated; work in progress already suggests its suitability for tropical forests (Schmitt et al., in prep). A further limitation, not intrinsic to the Fourier Transform itself but to how we applied it, concerns the definition of the sliding window used in the analysis. Clear trade-offs exist between window length, frequency resolution, and sensitivity to transient weather events, and these warrant further investigation in microclimate ecology and biogeography, even though they have already been extensively studied in applied mathematics [e.g. @richardson2018].

### **Perspectives**

Given these limited limitations, we believe the Fourier Transform approach nonetheless holds strong potential for microclimate ecology and biogeography. Future directions include: (i) further exploration of higher-order harmonics and harmonic hierarchies; (ii) benchmarking the Fourier Transform against alternative downscaling methods; and (iii) making the method accessible through open-source tools.

Higher-order harmonics such as sub-daily cycles and harmonic hierarchies can reveal fine-scale buffering mechanisms beyond the dominant 24-hour period, such as stomatal dynamics identified from latent and sensible heat fluxes in an Amazonian forest [@pinheiro2011], and thus deserve further exploration. Similarly, while we focused here on macroclimate debiasing using a simple mean-and-variance correction for comparison [@hes2025], more advanced techniques exist, including quantile mapping [@cannon2015] or machine learning [@coulibaly2005] calling for a systematic benchmarking of Fourier Transform debiasing against these alternatives.

Finally, we believe that wider adoption of the Fourier Transform approach in microclimate ecology and biogeography will require making the method accessible through open-source tools, such as an R package, a format increasingly favoured by ecologists [@lai2019]. This is a work in progress, with the active development of the R package *microclimr* and its associated application *microclimapp* (Schmitt et al., in prep), which also underlies the analyses presented in this paper (see Data availability).

### **Conclusion**

Taken together, our results show that the Fourier Transform is not merely an alternative to existing offset and slope-equilibrium approaches, but a methodological advance for microclimate ecology and biogeography. By resolving buffering across multiple timescales within a single, noise-robust framework, and by opening a path towards principled macroclimate debiasing, spectral decomposition addresses several limitations of current methods while remaining fully compatible with them. As microclimate data continues to be collected by sensor networks around the world, we believe that spectral approaches, such as the one presented here, and their future implementation in accessible, open-source tools, will become an increasingly valuable part of a microclimate ecologist's toolkit.

## References

::: {#refs}
:::

## Supplementary material

```{r buffer}
#| label: fig-buffer
#| message: false
#| warning: false
#| fig-cap: The amplitude ratio for the 24-hour period between May and September from 2020 to 2023 for air and soil sensors in the Mormal, Blois and Aigoual forests, for ONF and ERA5-Land macroclimate data.
buffers <- read_tsv("data/decomposition.tsv") %>%
  filter(period == 12) %>%
  select(source, site, plot, sensor, position, pai, power, datetime) %>%
  pivot_wider(names_from = source, values_from = power) %>%
  mutate(ratio_era = hobo / era, ratio_onf = hobo / onf) %>%
  select(-era, -hobo, -onf) %>%
  gather(source, ratio, -site, -plot, -sensor, -pai, -datetime, -position) %>%
  mutate(source = gsub("ratio_", "", source))
buffers %>%
  group_by(plot, position, source) %>%
  filter(n() > 107) %>%
  ungroup() %>%
  mutate(source = recode(source,
    "era" = "ERA5-Land",
    "onf" = "ONF weather station"
  )) %>%
  ggplot(aes(plot, ratio, fill = position, col = position)) +
  geom_violin() +
  facet_grid(source ~ site, scales = "free_x") +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  geom_hline(yintercept = 1) +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 14, face = "bold"),
    strip.text = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12)
  ) +
  ylab("Amplitude ratio for the period 24-hr [°C]") +
  scale_y_sqrt() +
  scale_fill_manual("", values = c("#a9d9e5", "#c89867")) +
  scale_color_manual("", values = c("#a9d9e5", "#c89867"))
```

```{r pai}
#| label: fig-pai
#| message: false
#| warning: false
#| fig-height: 8
#| fig-cap: The relationship between the mean amplitude ratio per sensor over the 24-hour period between May and September from 2020 to 2023 for air and soil sensors, and the plot area index in the Mormal, Blois and Aigoual forests.
buffers %>%
  group_by(plot, position, source) %>%
  filter(n() > 107) %>%
  ungroup() %>%
  group_by(source, site, plot, sensor, position, pai) %>%
  summarise(ratio = mean(ratio, na.rm = TRUE)) %>%
  mutate(source = recode(source,
    "era" = "ERA5-Land",
    "onf" = "ONF weather station"
  )) %>%
  ggplot(aes(pai, ratio, col = site)) +
  geom_point() +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  geom_smooth(method = "lm") +
  ggpubr::stat_regline_equation(
    aes(label = paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
    size = 4.5
  ) +
  theme(
    legend.position = "bottom",
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 12, face = "bold")
  ) +
  xlab("Plant Area Index [m2/m2]") +
  ylab("Mean Amplitude ratio for period 24-hr [°C]") +
  scale_color_discrete("") +
  facet_grid(position ~ source)
```

## Supplementary note

Analysis of time series and particularly periodic signal might be achieved with the help of the so-called Fourier Analysis and especifically for sampled signal with the Discrete Fourier Transform. These tools designed after the work of J. Fourier has been very fruitfull in a wide range of application to analyzed signal in the frequency domain.

Fourier Analysis is in fact a set of theoritical and computationnal tools, for which the most elementary is the so-called Fourier Series Decomposition we describe in the following. A simple periodic function like \$ A\cos(2\pi t/T+\varphi)\$ is a periodic function with period $T$; frequency $f = 1/T$ sometimes multiplied by $2\pi$ so that $\omega=2\pi f = 2\pi/T$ is called the pulsation; and its amplitude is $A$ because it ranges from its minimum $-A$ to its maximum $A$. Also the number $\varphi$ is known has the phase and only depends on the reference time $0$. The latter corresponds to a simple shift in time by $t_{0}=\varphi/(2\pi f)$ so that $\cos(2\pi t/T+\varphi)) = \cos(2\pi(t+t_{0})/T)$. The function has a mean $0$, thus a generic, $T$ periodic signal would be of the form $A\_{0} + A\_{1} \cos(2\pi t/T+\varphi\*{1})$ where $A_{0}$ is the mean, $A_{1}$ the amplitude and $\varphi_{1}$ the phase. However a general $T$-periodic function $S(t)$ is not has simple like this example, but a theorem due to Fourier, suggest under very few hypothesis (almost always true in application) any $T$- periodic function might be decomposed has a sum of its harmonic, says multiple of the fundamental frequency $f = 1/T$, namely $S(t) = A{0} + A\_{1} \cos(2\pi f t+\varphi\*{1}) + A*{2}* \cos(4\pi f t+\varphi\*{2}) + A{3} \cos(6 \pi f t+\varphi\_{3}) + \ldots$ Moreover, the numbers $A_{1},A_{2},\ldots$ decrease to $0$ significantly fast so that most of the signal information is contained in the first harmonics, that is the first coefficients $A_{0},A_{1},A_{2},\ldots$ and its corresponding phase $\varphi_{1},\varphi_{2},\ldots$\
Such a decomposition is called frequency or Fourier decomposition, this is the knowledge of two numbers $(A_k,\varphi_k)$ for each multiple of the fundamental frequency, says $f_{k}=kf$

It is worth noticing we have a one to one correspondence between a periodic signal and its Fourier decomposition, namely all the signal might be re-construct by virtue of the knowledge of $A_{0}, A_{1}, A_{2}, \cdots$ and the respective phase. The squared amplitudes $A_{0}^{2}, A_{1}^{2}, A_{2}^{2},
\cdots$ is called \emph{power spectrum} and a famous relation, the Parseval's Theorem relates the variance of the signal to the power spectrum by \[\frac 1 T \int\*{0}\^{T} \|S(t) -A*{0}\|\^{2} =* \frac 1 2 \sum\*{k=1}\^{\infty} A{k}\^{2} \] where $A_{0}=\frac 1 T \int_{0}^{T}S(t)dt$ is the mean. One of the strenght of the method is each coefficient $A_{k}$, is in fact an integration of the signal, see REF, this means that the coefficients represents and averaged quantities of the signal and not only a local (temporal) measure of the signal. This turns the method very robust to noisy signal due measurment error or local event of the captor.

Hypothetically, any signal is a periodic signal with infinite period thus having infinitely many frequency and phase, this goes to to a tool called the Fourier Transform, out of the scope of this presentation. But in application signal is analyzis on a finite time window $[0,T]$ where $T$ may range from days to years or decades depending on what matters (climatic, seasonnal, dayly effect). But this pieace of information can be seen as a portion of a periodic signal with period $T$ and the analysis follows as presented above.

As an example, microclimate data can be analyzed on few days window to avoid long trend effect (seasonal, climate change) and analyze the local response to a given condition of temperatures. Assume, climate data sampled at interval of 1 hour during 5 days, says $T=24\cdot 5$ and $N = 24\cdot 5$ data given by $S_{1},\ldots,S_{N}$: The best we can do is to approximate this points with the following Fourier series:

$$ S(t) = A\_{0} + \sum\*{k=1}\^{N/2} A\*{k}\cos ( 2k\pi f t + \varphi\_{k}) $$ with $A_{k}\geq 0$ for $k=1,\ldots,N/2$, and $f = 1/T$. The limitation to a multiple of the frequency $N/2$ is due to the fact that variation faster than the sampled rate cannot be captured. Note, the coefficient $A_{k}$ and $\varphi_{k}$ are computed with a very efficient method called the Fast Fourier Transform (FFT).

We notice the formula contains the multliple of the frequency $f =  1/5 \,d^{-1} = 1/120\, h^{-1}$. The first mutliple $f_{1}$, $f_{2}$ up to $f_{4}$ contains variation on the dayly to weekly range (meteorological variation), but this subset contains the frequency $f_{5} = 1/24 h^{-1}$ and its own multiple $f_{10}, f_{15}$, etc which contains the daily information and its harmonics. But temperatures signal exhibits clearly a dayly periodicity over a general trend, this is why comparing the powers between reference weather to microclimate air and soil decomposition is efficiently down by the comparison of the power spectrum for this peculiar frequencies.

Indeed, the principle of comparison of two sampled signal $S_{1},S_{2}, \ldots$ and $R_{1},R_{2}, \ldots$ at times $t_{1}, t_{2},\ldots$ (says macro and microclimate) may be done as a first approximation by the comparison of $S_n$ versus $R_n$, the method we proposed is to compare the Fourier coefficients $A_{k}(S)$ vs $B_{k}(R)$, respectively for the signal $S$ and $R$. The knowlege of the transformation of $A_{k}(S)$ to $A_{k}(R)$ by statistical approach allows to reconstruct a coherent signal $R$ from $S$ or the reverse. Of course we omit in this presentation the treatment of the phase which is in the same fashion but with the carefull attention that phase is not uniquely define (only modulo $2\pi k$). Maybe the most important fact in this representation is most of the transformation is contained in the part of the spectrum that capure most of the total power (see the Parseval Theorem). In our case it is generally dayly frequencies associate to the mean.

It is worth noticing that the method can be improved in particular with the use of windowing methods which is out of the scope of this presentation and more advanced methods from signal analyzis. But we believe this first treatment incorporates the key point of view to provide robust statistical analyzis of the transformation form macro to microclimate.

### Particular notes on the Fourier coeffcients.

For $N$ samples with $N$ even, the FFT methods (used in the package) render complex numbers $c_k= a_k-ib_k$ for $k=0,\ldots N/2$. Always, $b_0=0$ and $A_0 = c_0 = a_0$, the amplitudes for $k\geq 1$ are $$ A_k =  \sqrt{a_k^2 + b_k^2} $$ so the powers are $$ A_k^2 = a_k^2 + b_k^2 . $$

To compute the phase for $k\geq 1$ $$ \varphi_k = Arg(a_k-ib_k)$$ where $Arg$ is the function argument of a complex number.

The fourier series might be written equivalently

$$ S(t) = a_0 + \sum_{k=1}^{N/2} a_k \cos(2\pi k f t) + b_k \sin(2\pi k f t) $$
