---
title: "IMPRINT"
author: "Sylvain Schmitt, Erwan Hingant, Jonathan Lenoir"
date: last-modified
format: 
  html:
    embed-resources: true
    toc: true
    code-fold: true
bibliography: references.bib
---

```{r set}
#| message: false
# devtools::install_github("sylvainschmitt/miroclimr")
library(tidyverse)
library(dataverse)
library(microclimr)
```

## Data

We downloaded back the data from the IMPRINT project [@gril2024] using the INDORES dataverse and the R dataverse package. Additionally macroclimatic data were retrieved from ERA5-Land [@muñoz-sabater2021] using the Google Earth Engine (`data/get_era.py` script) or shared from ONF by Erwan Hingant (pers. com.).

```{r dataverse}
#| eval: false
plots <- list(
  Aigoual = "10.48579/PRO/RSBT3C/VDNSLZ",
  Blois = "10.48579/PRO/RSBT3C/37MKWL",
  Mormal = "10.48579/PRO/RSBT3C/IJAKFT"
) %>% lapply(get_dataframe_by_doi, server = "data.indores.fr") %>% 
  bind_rows(.id = "site")
write_tsv(plots, "data/plots.tsv")
plots %>% 
  group_by(site) %>% 
  summarise(lon = mean(longwgs84), lat = mean(latwgs84)) %>% 
  write_tsv("data/coords.tsv")
temperatures <- list(
  Aigoual = "10.48579/PRO/ZQBQUW/MK2F35",
  Blois = "10.48579/PRO/ZQBQUW/ZL0LOC",
  Mormal = "10.48579/PRO/ZQBQUW/GBVNRR"
) %>% lapply(get_dataframe_by_doi, server = "data.indores.fr") %>% 
  bind_rows(.id = "site")
write_tsv(temperatures, "data/hobo.tsv")
lidar <- list(
  Aigoual = "10.48579/PRO/XH62TC/P4F7FP",
  Blois = "10.48579/PRO/XH62TC/OI6WQU",
  Mormal = "10.48579/PRO/XH62TC/5RXUGU"
) %>% lapply(get_dataframe_by_doi, server = "data.indores.fr") %>% 
  bind_rows(.id = "site")
write_tsv(lidar, "data/lidar.tsv")
lidar_sub <- lidar %>% 
  filter(buffer == 5) %>% 
  rename(plot = idplot) %>% 
  select(site, plot, pai)
era <- read_tsv("data/era.tsv") %>% 
  mutate(site = recode(dim_0, "0" = "Aigoual", "1" = "Blois", "2" = "Mormal")) %>% 
  mutate(datetime = as_datetime(time)) %>% 
  rename(era = tas) %>% 
  select(site, datetime, era)
onf <- list.files("data/methor_data_onf/", pattern = ".txt", full.names = TRUE) %>% 
  vroom::vroom(col_names = read_tsv("data/methor_data_onf/MetHor2018.org") %>% 
                 names()) %>% 
  select("Code placette",
         "Date",
         "Heure (TU)",
         "Température instantanée (°C)") %>% 
  mutate(datetime = paste0(str_sub(Date, 7, 10), "-",
                           str_sub(Date, 4, 5), "-",
                           str_sub(Date, 1, 2), " ",
                           str_sub(`Heure (TU)`, 1, 2), ":",
                           str_sub(`Heure (TU)`, 3, 4), ":00")) %>% 
  mutate(datetime = as_datetime(datetime)) %>% 
  rename(onf = "Température instantanée (°C)", plot = "Code placette") %>% 
  filter(plot %in% c("CHP 59", "CHS 41", "HET 30")) %>% 
  mutate(site = recode(plot, "CHP 59" = "Mormal",
                       "CHS 41" = "Blois", "HET 30" = "Aigoual")) %>% 
  select(site, datetime, onf) %>% 
  mutate(onf = gsub(",", ".", onf)) %>% 
  mutate(onf = as.numeric(onf))
data <- temperatures %>% 
  mutate(datetime = as_datetime(datetime)) %>% 
  rename(hobo = t_hobo, plot = id_plot, sensor = id_sensor,
         position = position_sensor) %>% 
  mutate(position = recode(position, "a" = "air", "s" = "soil")) %>% 
  select(site, plot, sensor, position, datetime, hobo) %>% 
  left_join(era) %>% 
  left_join(onf) %>% 
  left_join(lidar_sub)
write_tsv(data, "data/data.tsv")
```

## Macroclimate

First, we compared the macroclimate data from ONF local stations with the ERA5-Land reanalysis. Despite an overestimation of low temperatures in ERA5-Land, especially in Aigoual, and a slight underestimation of very high temperatures in Mormal, they show high similarity.

```{r macroclimate}
#| message: false
#| warning: false
#| fig-cap: Macroclimatic temperature of ONF weather stations against ERA5-Land reanalysus. The black line represents the 1:1 line and the blue line a smoothing.
read_tsv("data/data.tsv") %>% 
  filter(month(datetime) %in% 5:9) %>% 
  ggplot(aes(onf, era)) +
  geom_point(alpha = .5, col = "lightgrey") +
  geom_abline() +
  geom_smooth(methode = "lm", se = FALSE) +
  facet_grid(position ~ site) +
  theme_bw() +
  xlab("ONF Temperature [°C]") +
  ylab("ERA5-Land Temperature [°C]")
```

## Decomposition

We decomposed all the temperature series for ERA5-Land and HOBOs in May to September for each plot and sensor on 5 days window sliding of 3 days using a single call to `microclimr::fft_roll` (ca. 2 minute):

```{r decomposition}
#| eval: false
data <- read_tsv("data/data.tsv")
fc <- data %>%
  filter(month(datetime) %in% 5:9) %>% 
  gather(source, temperature,
         -site, -plot, -sensor, -position, -datetime, -pai) %>% 
  group_by(source, site, plot, sensor, position, pai,
           datetime = floor_date(datetime, "2 hour")) %>%
  summarise(temperature = mean(temperature)) %>% 
    group_by(source, site, plot, sensor, position, pai) %>%
  do(fft = fft_roll(
    data = .,
    t = 5*12,
    index_col = "datetime",
    temperature_col = "temperature",
    window = "5 days",
    step = "3 days"
  )) %>%
  unnest(fft)
write_tsv(fc, "data/decomposition.tsv")
```

> We summarised the temperature data at a frequency of two hours to homogenise soil and air temperature sensors, but we could use different time frequencies in the future.

We then computed a semi-lognormal power spectrum for air temperatures, using HOBO microclimate temperatures and ERA5-Land and ONF macroclimate temperatures, for the locations of Aigoual, Blois and Mormal. The power spectrum revealed higher power at periods 0 and 24-hr for the ERA5-Land weather stations compared to the ONF weather stations, which could bias buffer estimates.

```{r power_spectrum}
#| message: false
#| warning: false
#| fig-cap: Semi-lognormal power spectrum for air temperatures, using HOBO microclimate temperatures and ERA5-Land and ONF macroclimate temperatures, for the locations of Aigoual, Blois and Mormal
"data/decomposition.tsv" %>% 
  read_tsv() %>% 
  filter(position == "air") %>% 
  group_by(source, site,  frequency, period) %>% 
  summarise(sd = sd(power, na.rm = TRUE),
            power = mean(power, na.rm = TRUE)) %>% 
  ggplot(aes(frequency, power+1)) +
  geom_col() +
  geom_linerange(aes(ymin = power+1-sd, ymax = power+1+sd)) +
  theme_bw() +
  xlab("Period [h]") +
  ylab("Power") +
  scale_x_continuous(
    breaks = c(0, 1 / 24*2, 1 / 12*2, 1 / 8*2, 1 / 6*2, 1 / 3*2),
    labels = c("0", "24", "12", "8", "6", "3")
  ) +
  theme(legend.position.inside = c(0.8, 0.8)) +
  facet_grid(site ~ source) +
  scale_y_log10()
```

## Buffer

We then computed the amplitude buffer backwards over a 24-hour period. Using ONF macroclimate data, we observed a significant negative temperature amplitude buffer at Blois and Mormal, but not with ERA5-Land macroclimate reanalysis.

```{r buffer}
#| message: false
#| warning: false
#| fig-cap: The amplitude ratio for the 24-hour period between May and September from 2020 to 2023 for air and soil sensors in the Mormal, Blois and Aigoual forests, for ONF and ERA5-Land macroclimate data.
buffers <- read_tsv("data/decomposition.tsv") %>% 
  filter(period == 12) %>% 
  select(source, site, plot, sensor, position, pai, power, datetime) %>% 
  pivot_wider(names_from = source, values_from = power) %>% 
  mutate(ratio_era = hobo/era, ratio_onf = hobo/onf) %>% 
  select(-era, -hobo, -onf) %>% 
  gather(source, ratio, -site, -plot, -sensor, -pai, -datetime, -position) %>% 
  mutate(source = gsub("ratio_", "", source))
buffers %>% 
  ggplot(aes(plot, ratio, fill = position, col = position)) +
  geom_violin() +
  facet_grid(source ~ site, scales = "free_x") +
  theme_bw() +
  scale_fill_discrete(guide = "none") +
  scale_color_discrete(guide = "none") +
  geom_hline(yintercept = 1) +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank()) +
  ylab("Amplitude ratio for the period 24-hr") +
  scale_y_sqrt()
```

```{r buffer_summary}
#| message: false
#| warning: false
#| fig-cap: The mean amplitude ratio per sensor for the 24-hour period between May and September from 2020 to 2023 for air and soil sensors in the Mormal, Blois and Aigoual forests, for ONF and ERA5-Land macroclimate data.
buffers %>% 
  ggplot(aes(position, ratio, fill = site)) +
  geom_boxplot() +
  theme_bw() +
  coord_flip() +
  ylab("Mean amplitude ratio per sensor for the period 24-hr") +
  xlab("") +
  scale_y_sqrt() +
  facet_wrap(~ source)
```

## Vegetation

Then we can easily establish the relationship between the plant area index and the air temperature in Blois and Mormal.

```{r pai}
#| message: false
#| warning: false
#| fig-height: 8
#| fig-cap: The relationship between the mean amplitude ratio per sensor over the 24-hour period between May and September from 2020 to 2023 for air and soil sensors, and the plot area index in the Mormal, Blois and Aigoual forests.
buffers %>% 
  group_by(source, site, plot, sensor, position, pai) %>% 
  summarise(ratio = mean(ratio, na.rm = TRUE)) %>% 
  ggplot(aes(pai, ratio, col = site)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method = "lm") +
  ggpubr::stat_regline_equation(
    aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
  ) +
  theme(legend.position = "bottom") +
  xlab("Plant Area Index") +
  ylab("Mean Amplitude ratio for period 24-hr") +
  scale_color_discrete("Site") +
  facet_grid(source ~ position)
```
